---
title: "Genereate_Compare_output"
author: "Torda"
date: "February 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(phytools)
library(ape)
library(truncnorm)
library(stringr)
library(readr)
library(plyr)
```

A lenyeg hogy le kene generalni kulonbozo meretu fakat kulonbozo meretu cluster nagysagokkal

pl 10, 100, 1000, 10000 fajszammal

aztan 100, 500, 1000 clusterrel

a clusterek ortholog csoportjai pedig egy tuncated normal eloszlas mellett kene kiosztani

```{r}
trees <- pbtree(n = 100)
is.ultrametric(trees)
tree <- rtree(100)
is.ultrametric(tree)
plot(tree)
is.rooted(tree)
trees <- rmtree(10, 100)
```

```{r}
n <- rnorm(n = 100, mean = 50, sd = 10)
hist(round(n))
min(n)
n <- rtruncnorm(100, a = 0, mean = 50, sd = 40)
hist(round(n))

length(sample(1:1000, size = 20, prob = 1- pexp(1:1000)))
hist((1- pexp(1:1000, rate = 0.1)))
hist(round(sample(1:1000, size = 10000, replace = TRUE, prob = 1- pexp(1:1000, rate = 0.01))))
any(round(sample(1:1000, size = 10000, replace = TRUE, prob = 1- pexp(1:1000, rate = 0.01))) == 0)
min(round(sample(1:1000000, size = 100000, replace = TRUE, prob = 1- pexp(1:1000000, rate = 0.01))))
max(round(sample(1:1000000, size = 100000, replace = TRUE, prob = 1- pexp(1:1000000, rate = 0.00001))))
max(round(sample(1:1000, size = 10000, replace = TRUE, prob = 1- pexp(1:1000, rate = 0.00001))))
hist(round(sample(1:1000, size = 10000, replace = TRUE, prob = 1- pexp(1:1000, rate = 0.00001))))

1 / 1000000
0.00001

plot_ortho_hist <- function(ortho_n, cl_num){
  hist(round(sample(1:ortho_n, size = cl_num, replace = TRUE, prob = 1- pexp(1:ortho_n, rate = 1 / (ortho_n / 10)))))
}
plot_ortho_hist(10000, 100)
plot_ortho_hist(100000, 100)
plot_ortho_hist(1000000, 100)
```

generateing tree file and dolloout species iden

```{r}
spec_num <- 10
tree <- rtree(spec_num)
nodes <- tree$Nnode + spec_num
species_df <- data.frame(nodes = 1:nodes, species = c(tree$tip.label, sapply((spec_num + 1):nodes, function(x) str_c(extract.clade(tree, x)$tip.label, collapse = " "))), stringsAsFactors = FALSE)
sink(file = "probe.txt")
cat("For your reference, here is the list of internal nodes of the tree and the descendants for each\n")
for(i in 1:nrow(species_df)){
  cat(paste0("node", species_df[i,1]), "\t ", species_df[i,2], "\n", sep = "")
}
sink()

```


generate cluster / ortholog data frame

```{r}
ortho_dist <- function(ortho_n, cl_num){
  round(sample(1:ortho_n, size = cl_num, replace = TRUE, prob = 1- pexp(1:ortho_n, rate = 1 / (ortho_n / 1000))))
}


cl_num <- 10000000
ortho_upper <- 1000000
n <- ortho_dist(ortho_n = ortho_upper, cl_num = cl_num)
cl_df <- data.frame(clusters = 1:cl_num, ortho_n = n)
sum(cl_df$ortho_n)
max(cl_df$ortho_n)
```

check a real data,

```{r}
library(CompareTools)
```

```{r}
eni_tree <- read.tree("~/Documents/MUNKA/PhD/CompareTools/inst/extdata/71g_species.tree")
files <- "~/Documents/MUNKA/PhD/CompareTools/inst/extdata/Dolloout_GW_all"  
comp <- DolloImport2(tree = eni_tree, files = files, partitioning = FALSE, part_name = "all")
df <- comp$all$all_cluster
sum(df[,2]) / sum(df[,1])
sum(df[,1])
comp2 <- DolloImport2(tree = eni_tree, files = files, partitioning = TRUE, part_name = "every")
length(comp2$every)
gain <- sapply(comp2$every, function(x) sum(x[,1]))
hist(gain)
summary(gain)
```

tehat 144961 cluster eseten
max 49541 otho van egy clusterban, min 1, median 1. mean 13.18 3rd Qu. 2.


